### 2018-06-13 express2018/context.tex
Minor fixes

A \emph{weak guarded} context is a context in which all holes appears
in the sub-expression of prefixed processes. We can easily define it
following the same idea:
\begin{alltt}
\HOLConst{WG} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{p})
\HOLConst{CONTEXT} \HOLFreeVar{e} \HOLSymConst{\HOLTokenImp{}} \HOLConst{WG} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{a}\HOLSymConst{..}\HOLFreeVar{e} \HOLBoundVar{t})
\HOLConst{WG} \HOLFreeVar{e\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{WG} \HOLFreeVar{e\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{WG} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{e\sb{\mathrm{1}}} \HOLBoundVar{t} \HOLSymConst{+} \HOLFreeVar{e\sb{\mathrm{2}}} \HOLBoundVar{t})
\HOLConst{WG} \HOLFreeVar{e\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{WG} \HOLFreeVar{e\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{WG} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{e\sb{\mathrm{1}}} \HOLBoundVar{t} \HOLSymConst{\ensuremath{\parallel}} \HOLFreeVar{e\sb{\mathrm{2}}} \HOLBoundVar{t})
\HOLConst{WG} \HOLFreeVar{e} \HOLSymConst{\HOLTokenImp{}} \HOLConst{WG} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLSymConst{\ensuremath{\nu}} \HOLFreeVar{L} (\HOLFreeVar{e} \HOLBoundVar{t}))
\HOLConst{WG} \HOLFreeVar{e} \HOLSymConst{\HOLTokenImp{}} \HOLConst{WG} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLConst{relab} (\HOLFreeVar{e} \HOLBoundVar{t}) \HOLFreeVar{rf})\hfill{[WG_rules]}
\end{alltt}
(Notice the differences between a weak guarded context and a normal
one: $\lambda t. t$ is not weakly guarded as the variable is directly
exposed without any prefixed action. And $\lambda t. a.e[t]$ is weakly
guarded as long as $e[\cdot]$ is a  context, no necessary weakly guaded.)

It's a similar process to define strongly guarded context
(\texttt{SG}), sequential context (\texttt{SEQ}) and their
variants without direct sums (\texttt{GCONTEXT}, \texttt{WGS},
\texttt{GSEQ}). Some lemmas about their relationships have very long
(but trivial) formal proofs due to the combinatorial explosion of
inductions on their structures. For example, one such 
lemma says, for any  context $E$ which is both strongly guarded
and sequential (no direct sums) and another sequential context $H$ (no
direct sums), the composition $H \circ E$ is still both strongly
guarded and sequential (no direct sums):
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E}. \HOLConst{SG} \HOLBoundVar{E} \HOLSymConst{\HOLTokenConj{}} \HOLConst{GSEQ} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{H}. \HOLConst{GSEQ} \HOLBoundVar{H} \HOLSymConst{\HOLTokenImp{}} \HOLConst{SG} (\HOLBoundVar{H} \HOLSymConst{\HOLTokenCompose} \HOLBoundVar{E}) \HOLSymConst{\HOLTokenConj{}} \HOLConst{GSEQ} (\HOLBoundVar{H} \HOLSymConst{\HOLTokenCompose} \HOLBoundVar{E})
\end{alltt}

>>>>

A \emph{weak guarded} context is a context in which all holes appears
in the sub-expression of prefixed processes. We can easily define it
following the same idea:
\begin{alltt}
\HOLConst{WG} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{p})
\HOLConst{CONTEXT} \HOLFreeVar{e} \HOLSymConst{\HOLTokenImp{}} \HOLConst{WG} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{a}\HOLSymConst{..}\HOLFreeVar{e} \HOLBoundVar{t})
\HOLConst{WG} \HOLFreeVar{e\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{WG} \HOLFreeVar{e\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{WG} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{e\sb{\mathrm{1}}} \HOLBoundVar{t} \HOLSymConst{+} \HOLFreeVar{e\sb{\mathrm{2}}} \HOLBoundVar{t})
\HOLConst{WG} \HOLFreeVar{e\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{WG} \HOLFreeVar{e\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{WG} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{e\sb{\mathrm{1}}} \HOLBoundVar{t} \HOLSymConst{\ensuremath{\parallel}} \HOLFreeVar{e\sb{\mathrm{2}}} \HOLBoundVar{t})
\HOLConst{WG} \HOLFreeVar{e} \HOLSymConst{\HOLTokenImp{}} \HOLConst{WG} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLSymConst{\ensuremath{\nu}} \HOLFreeVar{L} (\HOLFreeVar{e} \HOLBoundVar{t}))
\HOLConst{WG} \HOLFreeVar{e} \HOLSymConst{\HOLTokenImp{}} \HOLConst{WG} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLConst{relab} (\HOLFreeVar{e} \HOLBoundVar{t}) \HOLFreeVar{rf})\hfill{[WG_rules]}
\end{alltt}
(Notice the differences between a weak guarded context and a normal
one: $\lambda t. t$ is not weakly guarded as the variable is directly
exposed without any prefixed action. And $\lambda t. a.e[t]$ is weakly
guarded as long as $e[\cdot]$ is a context, not necessary weakly guaded.)

It's a similar process to define strongly guarded context
(\texttt{SG}), sequential context (\texttt{SEQ}) and their
variants without direct sums (\texttt{GCONTEXT}, \texttt{WGS},
\texttt{GSEQ}). Some lemmas about their relationships have very long
(but trivial) formal proofs due to multiple levels of
inductions on context structures. For example, one such 
lemma says, for any  context $E$ which is both strongly guarded
and sequential (no direct sums) and another sequential context $H$ (no
direct sums), the composition $H \circ E$ is still both strongly
guarded and sequential (no direct sums):
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E}. \HOLConst{SG} \HOLBoundVar{E} \HOLSymConst{\HOLTokenConj{}} \HOLConst{GSEQ} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{H}. \HOLConst{GSEQ} \HOLBoundVar{H} \HOLSymConst{\HOLTokenImp{}} \HOLConst{SG} (\HOLBoundVar{H} \HOLSymConst{\HOLTokenCompose} \HOLBoundVar{E}) \HOLSymConst{\HOLTokenConj{}} \HOLConst{GSEQ} (\HOLBoundVar{H} \HOLSymConst{\HOLTokenCompose} \HOLBoundVar{E})
\end{alltt}

### 2018-06-13 express2018/contraction.tex
Added missing equa.tex
Minor fixes

\section{Contractions}
\label{s:mcontr}
The constraints on the unique-solution Theorem~\ref{t:Mil89} can be 
weakened if  we move from equations to inequations called
  \emph{contractions}.

@@

\item 
    $Q \arr\mu Q'$   implies there is $P'$ such that $P \Arcap \mu
 P'$ and $P' 
\wb Q'$.
\end{enumerate}
\emph{Bisimilarity  contraction}, written $\mcontrBIS$, is the union
of all bisimulation contractions. 
\qed\end{definition}
% We say that $P$  {\emph{bisimilarly  contracts} $Q$, written
% $Q \mcontrBIS Q$.
% We write $\mexpaBIS$ for the inverse of $\mcontrBIS$.

@@

Bisimilarity  contraction is coarser than 
 the expansion  relation 
$\contr$ of   Definition~\ref{d:expa}. Clause (1) is the same in the two
definitions. But in clause (2) expansion uses 
$P \Arr \mu P'$, rather than $P \Arcap \mu P'$; 
 moreover with
contraction the final derivatives are simply required to be bisimilar.
An expansion 
$P \contr Q$
tells us  that $Q$ is always at least as efficient as $P$, whereas  the
 contraction $P \mcontrBIS Q$  just says that $Q$ has the  possibility of
being at least as efficient as $P$.

@@

\begin{example}
\label{exa:contr}
{ We have %\mcontrBIS a + \tau^n . a $
 $ a \not  \mcontrBIS \tau. a$. However,
$a+ \tau . a \mcontrBIS a$, as well as its converse, 
$  a \mcontrBIS a +
\tau . a $. Indeed, if $P \wb Q$ then 
$  P  \mcontrBIS P +Q$. The last two relations do not hold with 
$\contr$, which explains the strictness of the inclusion
 ${\contr} \subseteq {\mcontrBIS}$. 
% The inclusion is strict: for instance
% $a+ \tau . a \mcontrBIS a$, where $\mcontrBIS$ cannot be replaced by
%  $\contr$. Also the converse of  $a+ \tau . a \mcontrBIS a$ holds, namely
% $  a \mcontrBIS a +
% \tau . a $. However, we have %\mcontrBIS a + \tau^n . a $
%  $ a \not  \mcontrBIS \tau. a$
\qed} \end{example}

>>>>

\section{Contractions}
\label{s:mcontr}

The constraints on the unique-solution Theorem~\ref{t:Mil89} can be 
weakened if  we move from equations to inequations called
  \emph{contractions}.

@@

\item 
    $Q \arr\mu Q'$   implies there is $P'$ such that $P \Arcap \mu
 P'$ and $P' 
\wb Q'$.
\end{enumerate}
\emph{Bisimilarity  contraction}, written $\mcontrBIS$, is the union
of all bisimulation contractions. 
\end{definition}
% We say that $P$  {\emph{bisimilarly  contracts} $Q$, written
% $Q \mcontrBIS Q$.
% We write $\mexpaBIS$ for the inverse of $\mcontrBIS$.

@@

Bisimilarity  contraction is coarser than 
 the expansion  relation 
$\expa$ of Definition~\ref{d:expa}. Clause (1) is the same in the two
definitions. But in clause (2) expansion uses 
$P \Arr \mu P'$, rather than $P \Arcap \mu P'$; 
 moreover with
contraction the final derivatives are simply required to be bisimilar.
An expansion 
$P \contr Q$
tells us  that $Q$ is always at least as efficient as $P$, whereas  the
 contraction $P \mcontrBIS Q$  just says that $Q$ has the  possibility of
being at least as efficient as $P$.

@@

\begin{example}
\label{exa:contr}
We have %\mcontrBIS a + \tau^n . a $
 $ a \not  \mcontrBIS \tau. a$. However,
$a+ \tau . a \mcontrBIS a$, as well as its converse, 
$  a \mcontrBIS a +
\tau . a $. Indeed, if $P \wb Q$ then 
$  P  \mcontrBIS P +Q$. The last two relations do not hold with 
$\expa$, which explains the strictness of the inclusion
 ${\expa} \subseteq {\mcontrBIS}$. 
% The inclusion is strict: for instance
% $a+ \tau . a \mcontrBIS a$, where $\mcontrBIS$ cannot be replaced by
%  $\contr$. Also the converse of  $a+ \tau . a \mcontrBIS a$ holds, namely
% $  a \mcontrBIS a +
% \tau . a $. However, we have %\mcontrBIS a + \tau^n . a $
%  $ a \not  \mcontrBIS \tau. a$
\end{example}

### 2018-06-13 express2018/formal.tex
Minor fixes

We highlight here a formalisation of CCS
in the HOL theorem
prover (HOL4) \cite{slind2008brief},
including the new concepts and theorems proposed in the first half of
this paper.
%  The main purpose is to convince the readers that, there's no flaw
%  in the informal proofs. 
The whole formalisation can be found 
in \cite{Tian:2017wrba}  or in 
 in the official example folder of HOL4 source
code\footnote{\url{https://github.com/HOL-Theorem-Prover/HOL}}. The
work consists of about 20,000 lines of proof scripts in Standard ML.
%  This
%  work is based on a formalisation of CCS in HOL theorem
% prover (HOL4) \cite{slind2008brief}, described in second author's MSc thesis
% \cite{Tian:2017wrba} and now lives in the official examples folder of HOL4 source
% code\footnote{\url{https://github.com/HOL-Theorem-Prover/HOL}}.

>>>>

We highlight here a formalisation of CCS
in the HOL theorem
prover (HOL4) \cite{slind2008brief},
including the new concepts and theorems proposed in the first half of
this paper.
%  The main purpose is to convince the readers that, there's no flaw
%  in the informal proofs. 
The whole formalisation can be found 
in \cite{Tian:2017wrba}  or in 
 in the official example folder of HOL4 source
code\footnote{\url{https://github.com/HOL-Theorem-Prover/HOL}}. The
work consists of about 20,000 lines of proof scripts in Standard ML,
derived from an early work \cite{Nesi:1992ve} by Monica Nesi in 1992-1995.

HOL4 is written in Standard ML (or ML for abbreviation). This single language plays three roles here:
\begin{enumerate}
\item It serves as the underlying implementation language for the core HOL engine;
\item It's used to implement tactics (and tacticals) for automatic theorem proving;
\item It's used as the command language of the interactive HOL system.
\end{enumerate}
Besides, HOL4 users can write arbitrary applications which leverages
the theorem proving facility.

### 2018-06-13 express2018/intro.tex
Minor fixes

The purpose of this paper is twofold. 
On the one hand, we refine the contraction proof technique by moving to the 
\emph{rooted contraction} \cite{Tian:2017wrba}, that is, the coarsest (pre)congruence contained in the contraction
preorder.  Using rooted contraction one obtains an unique-solution theorem that is valid for
\emph{rooted bisimilarity} (hence also for bisimilarity itself).  
On the other hand, we provide a  
 comprehensive formalisation  of the core of the theory of CCS 
 in the HOL
theorem prover (HOL4). The formalisation  includes the proofs of
Milner's 3 ``unique solution of equations'' theorems and
contractions discussed in the present paper, but is not limited to it (partly because such
theorems rely on a number of more fundamental results):
indeed the formalisation encompasses the basic properties of strong and weak
bisimilarity (e.g. the fixed-point and substitutivity properties), 
their algebraic theory, various versions of ``bisimulation up to''
techniques (e.g., bisimulation up-to expansion),
the basic properties  of rooted bisimilarity. Considering the relationship between bisimilarity and rooted
bisimilarity, the formalisation includes the proof that the latter is the coarsest
congruence included in the former, for which two proofs are formalized: one as in Milner's
book,  requiring the hypothesis that  no processes can use all labels; the other without
such hypothesis, essentially formalising van Glabbek's paper \cite{vanGlabbeek:2005ur}.
Similar theorems are proved for rooted contractions wrt the contraction preorder.

>>>>

The purpose of this paper is twofold. 
On the one hand, we refine the contraction proof technique by moving to the 
\emph{rooted contraction} \cite{Tian:2017wrba}, that is, the coarsest (pre)congruence contained in the contraction
preorder.  Using rooted contraction one obtains an unique-solution theorem that is valid for
\emph{rooted bisimilarity} (hence also for bisimilarity itself).  
On the other hand, we provide a  
 comprehensive formalisation  of the core of the theory of CCS 
 in the HOL
theorem prover (HOL4). The formalisation  includes the proofs of
``unique solution of (rooted) contractions'' theorems and
(pre)congruence properties discussed in the present paper, but is not limited to it (partly because such
theorems rely on a number of more fundamental results):
indeed the formalisation encompasses the basic properties of strong and weak
bisimilarity (e.g. the fixed-point and substitutivity properties), 
their algebraic theory, various versions of ``bisimulation up to''
techniques (e.g., bisimulation up-to expansion),
some basic properties of rooted bisimilarity (including Hennessy Lemma
and Deng Lemma).
Considering the relationship between bisimilarity and rooted
bisimilarity, the formalisation includes the proof that the latter is the coarsest
congruence included in the former, for which two proofs are formalized: one as in Milner's
book,  requiring the hypothesis that  no processes can use all labels; the other without
such hypothesis, essentially formalising van Glabbek's paper \cite{vanGlabbeek:2005ur}.
Similar theorems are proved for rooted contractions wrt the contraction preorder.

### 2018-06-13 express2018/part2.tex
Minor fixes

If there's at most one visible action (label) in the list of actions of a trace,
then the trace is also a weak transition. We divided this observation
into two cases: no label and unique label. The definition of ``no
label'' in an action list is easy and clear (\texttt{MEM} is for
testing if an element is \emph{member} of a list):
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{NO_LABEL} \HOLFreeVar{L} \HOLSymConst{\HOLTokenEquiv{}} \HOLSymConst{\HOLTokenNeg{}}\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{l}. \HOLConst{MEM} (\HOLConst{label} \HOLBoundVar{l}) \HOLFreeVar{L}\hfill{[NO_LABEL_def]}
\end{alltt}
while the definition of ``unique label'' can be done in many ways, in
which we have chosen to use the version learnt from Robert Beers in
a private discussion which prevented counting or filtering in the list:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{UNIQUE_LABEL} \HOLFreeVar{u} \HOLFreeVar{L} \HOLSymConst{\HOLTokenEquiv{}}
   \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{L\sb{\mathrm{1}}} \HOLBoundVar{L\sb{\mathrm{2}}}. (\HOLBoundVar{L\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenDoublePlus} [\HOLFreeVar{u}] \HOLSymConst{\HOLTokenDoublePlus} \HOLBoundVar{L\sb{\mathrm{2}}} \HOLSymConst{=} \HOLFreeVar{L}) \HOLSymConst{\HOLTokenConj{}} \HOLConst{NO_LABEL} \HOLBoundVar{L\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{NO_LABEL} \HOLBoundVar{L\sb{\mathrm{2}}}\hfill{UNIQUE_LABEL_def}
\end{alltt}
It says, a label is unique in an action list iff it there's no other
labels in the rest part of the list.

@@

The proof of Unique solution of rooted contractions theorem is the
same proof steps of Unique solution of contractions theorem plus the
application of \texttt{OBS_CONGR_BY_WEAK_BISIM} at the
beginning.\footnote{In the thesis work, the conclusion of this theorem
is weak bisimilarity, and the proof is exactly the same as Unique
solution of contractions theorem. Now we got a stronger conclusion
using observational congruence, and the previous version becomes a
trivial collorary of the current version.} The
two proofs are quite similar, mostly because the only property we need
from (rooted) contraction is its precongruence. Once we have proved
the precongruence of rooted contracion, we can naturally use the
normal version of weakly guarded expressions with direct sums included.

>>>>

If there's at most one visible action (label) in the list of actions of a trace,
then the trace is also a weak transition. We divided this observation
into two cases: no label and unique label. The definition of ``no
label'' in an action list is easy and clear (\texttt{MEM} is for
testing if an element is \emph{member} of a list):
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{NO_LABEL} \HOLFreeVar{L} \HOLSymConst{\HOLTokenEquiv{}} \HOLSymConst{\HOLTokenNeg{}}\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{l}. \HOLConst{MEM} (\HOLConst{label} \HOLBoundVar{l}) \HOLFreeVar{L}\hfill{[NO_LABEL_def]}
\end{alltt}
while the definition of ``unique label'' can be done in many ways, in
which we have chosen to use the version learnt from Robert Beers in
a private discussion which prevented counting or filtering in the list:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{UNIQUE_LABEL} \HOLFreeVar{u} \HOLFreeVar{L} \HOLSymConst{\HOLTokenEquiv{}}
   \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{L\sb{\mathrm{1}}} \HOLBoundVar{L\sb{\mathrm{2}}}. (\HOLBoundVar{L\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenDoublePlus} [\HOLFreeVar{u}] \HOLSymConst{\HOLTokenDoublePlus} \HOLBoundVar{L\sb{\mathrm{2}}} \HOLSymConst{=} \HOLFreeVar{L}) \HOLSymConst{\HOLTokenConj{}} \HOLConst{NO_LABEL} \HOLBoundVar{L\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{NO_LABEL} \HOLBoundVar{L\sb{\mathrm{2}}}\hfill{[UNIQUE_LABEL_def]}
\end{alltt}
It says, a label is unique in an action list iff it there's no other
labels in the rest part of the list.

@@

The formal proof of ``unique solution of rooted contractions theorem''
(Thm.\; \ref{t:rcontraBisimulationU}) has the
same initial proof steps as Thm.\; ref{{t:contraBisimulationU}} plus a
little more steps to handle the rooted bisimilarity at conclusion. (In the thesis work, the conclusion of this theorem
is weak bisimilarity, and the proof is exactly the same as Unique
solution of contractions theorem. Now we got a stronger conclusion
using observational congruence, and the previous version becomes a
trivial collorary of the current version.) The
two proofs are quite similar, mostly because the only property we need
from (rooted) contraction is its precongruence. Once we have proved
the precongruence of rooted contracion, we can naturally use the
normal version of weakly guarded expressions with direct sums included.

### 2018-06-13 express2018/sos.tex
DS
Minor fixes

HOL4 is written in Standard ML (or ML for abbreviation). This single language plays three roles here:
\begin{enumerate}
\item It serves as the underlying implementation language for the core HOL engine;
\item It's used to implement tactics (and tacticals) for automatic theorem proving;
\item It's used as the command language of the interactive HOL system.
\end{enumerate}
Besides, HOL4 users can write arbitrary applications which leverages
the theorem proving facility. The ultimate result (initially done by Nesi)
at this point, is a complex Standard ML function
  taking a CCS process and returning a theorem indicating all its direct transitions.

>>>>

The ultimate result (initially done by Nesi)
at this point, is a complex Standard ML function
  taking a CCS process and returning a theorem indicating all its direct transitions.

### 2018-06-13 express2018/unique.tex
Minor fixes

\begin{definition}
\label{d:uniContra}
Given a  system of contractions 
$\{  X_i \mcontr E_i\}_{i\in I}$, 
 we say that:
\begin{itemize}
\item
 $\til P$ is a \emph{solution  for $\mcontrBIS$ of the 
 system of contractions} 
 if $\til P \mcontrBIS \til E [\til P]$;
\item 
the system  
has  \emph{a unique 
solution for $\approx$}
if 
whenever 
 $\til P$ and $\til Q$ are both solutions  for  $\mcontrBIS$
 then $\til P \approx \til Q$.
\qed\end{itemize}
\end{definition}

@@

The system use \emph{weakly-guarded sums} if 
each $E_i$ only makes use of guarded sums.
\qed\end{definition}

@@

The unique solution theorem of Section~\ref{ss:SysContr} require a
constrained syntax for sums, due to the congruence and precongruence
problems of bisimilarity and contraction with such operator. 
We show here that the constraints can be
removed by moving to the induced congruence and precongruence.

\begin{definition}[rooted contraction]
\label{d:rcontra}
Two processes $P$ and $Q$ are in the \emph{rooted contraction}, written 
 $P\rcontr Q$, if
\begin{enumerate}
\item $P \arr\mu P'$ implies that there is $Q'$ with $Q \arr \mu Q'$
 and $P'\mcontrBIS Q'$;
\item $Q \arr\mu Q'$   implies that there is $P'$ with $P \Arr \mu
 P'$ and $P' \wb Q'$.
\end{enumerate}
\end{definition}

\begin{theorem}
\label{t:rcontrPrecongruence}
Relation $\rcontr$ is a precongruence in CCS.  Indeed it is the
largest precongruence contained in $\contr$.
\end{theorem}

The proof of this result is along the lines of the analogous result
for rooted bisimilarity with respect to bisimilarity. 
Similarly one can define and handle the rooted variant of the expansion preorder.

@@

\begin{theorem}[unique solution of contractions for $\rapprox$]
\label{t:rcontraBisimulationU}
A system of weakly-guarded contractions
    has 
a unique solution 
 for $\rapprox$.
\end{theorem}

>>>>

\begin{definition}
\label{d:uniContra}
Given a  system of contractions 
$\{  X_i \mcontr E_i\}_{i\in I}$, 
 we say that:
\begin{itemize}
\item
 $\til P$ is a \emph{solution  for $\mcontrBIS$ of the 
 system of contractions} 
 if $\til P \mcontrBIS \til E [\til P]$;
\item 
the system  
has  \emph{a unique 
solution for $\approx$}
if 
whenever 
 $\til P$ and $\til Q$ are both solutions  for  $\mcontrBIS$
 then $\til P \approx \til Q$.
\end{itemize}
\end{definition}

@@

The system use \emph{weakly-guarded sums} if 
each $E_i$ only makes use of guarded sums.
\end{definition}

@@

The unique solution theorem of Section~\ref{ss:SysContr} require a
constrained syntax for sums, due to the congruence and precongruence
problems of bisimilarity and contraction with such operator. 
We show here that the constraints can be
removed by moving to the induced congruence and precongruence, so
called ``rooted contraction'':
\begin{definition}[rooted contraction]
\label{d:rcontra}
Two processes $P$ and $Q$ are in the \emph{rooted contraction}, written 
 $P\rcontr Q$, if
\begin{enumerate}
\item $P \arr\mu P'$ implies that there is $Q'$ with $Q \arr \mu Q'$
 and $P'\mcontrBIS Q'$;
\item $Q \arr\mu Q'$   implies that there is $P'$ with $P \Arr \mu
 P'$ and $P' \wb Q'$.
\end{enumerate}
\end{definition}

The quick finding of rooted contraction was by help of
theorem provers, and was based on two principles: 1) Its definition must NOT be recursive
(or coinductive), instead it should mimic the definition of rooted bisimilarity
($\approx^c$); 2) It must be built on top of the existing \emph{contracts}
relation $\mcontrBIS$, which we believe it's the \emph{right} one (as
it's complete). Multiple candicates were tested, and finally the above
definition was proven to be a precongruence:
\begin{theorem}
\label{t:rcontrPrecongruence}
Relation $\rcontr$ is a precongruence in CCS.  Indeed it is the
largest precongruence contained in $\contr$.
\end{theorem}  
The proof of this result is along the lines of the analogous result
for rooted bisimilarity with respect to bisimilarity. 
Similarly one can define and handle the rooted variant of the expansion preorder.

@@

\begin{theorem}[Unique solution of contractions for $\rapprox$]
\label{t:rcontraBisimulationU}
A system of weakly-guarded contractions
    has 
a unique solution 
 for $\rapprox$.
\end{theorem}
